---
title: "Threatened Australians: A model for catalysing conservation knowledge and action through a digital platform"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
bibliography: /home/gareth/science/admin/bibliography/zotero_lib_exports/lenovo_pop_os_my_library_2.bib
biblio-style: apsr
link-citations: yes
csl: /home/gareth/science/admin/bibliography/csls/apa.csl
output:
  word_document:
    reference_docx: /home/gareth/science/admin/bibliography/word_templates/research_article_template.docx
    fig_capton: true
editor_options:
  markdown:
    wrap: 72
---

## Authors

Gareth S. Kindler^1,^\*, Nick Kelly^2^, James E.M. Watson^2^, social
scientist?, digital tech/advertising/GA4 wiz? (more to help me), Ariadne
Vromen (writing some impressive-looking book), Mitchell Whitelaw (), app
designer tech head, ]

## Affiliations

1.  School of Earth and Environmental Sciences, The University of
    Queensland, S t Lucia 4072, Australia
2.  [Add your affil here please Nick] School, Queensland University of
    Technology, suburb, postcode, Australia

\***Corresponding author email**:
[g.kindler\@uq.edu.au](mailto:g.kindler@uq.edu.au){.email}

**Keywords**: threatened species, digital technology, conservation
science, design

Target journals: Any of conservation journal (?), Journal of
Environmental Media
(<https://www.intellectbooks.com/journal-of-environmental-media>)

\newpage

<!-- ## Model papers -->

<!-- The papers I have found are about exploratory/educational/interactive -->

<!-- tools that visualise some aspect of the author's work. That's not -->

<!-- completely dissimilar to TA but it's not quite where we are at. I guess -->

<!-- people/orgs who develop advocacy web apps like TA don't write about it -->

<!-- in journals? -->

<!-- -   VolcanNoseR - <https://doi.org/10.1038/s41598-020-76603-3> -->

<!-- -   SpeciesMap - <https://doi.org/10.1007/s10531-012-0284-4> -->

<!-- -   BioDem - <https://doi.org/10.1111/jbi.14256> -->

<!-- -   Pop viability - <https://doi.org/10.1111/nrm.12120> -->

<!-- -   Niggle - <https://doi.org/10.2196/21085> -->

# Draft

## Abstract

Government action for species conservation is dependent upon public
support. Whilst scientists, NGOs, and government have developed tools
that improve accessibility, they fall short of engaging the broader
public. Here we discuss the Threatened Australians (TA) web app, which
uses existing data to tell the story of threatened species occurrence
within government constituency boundaries, as an attempt to facilitate
public support and thus political action. By examining the development,
launch, and maintenance, this helps inform/improve future efforts the
conservation community make in engaging broader audiences.

## Introduction

### Para 1 - Biodiversity crisis in Australia

Keys refs: Kearney, Ward, State of Environments etc

Challenge for humanity [IPBES]. Ecosystems and species are entangled in
'Biodiveristy', both of which are disappear at an undesirable rate.
Australia's biodiversity continues to decline due to human-led impacts.
Ecosystems are an integral part of the Australian natural capital, with
humans valuing in many different ways - relational, intrinsic too.
Despite being relatively well-off, Australia continues to allow its
endemic species to decline and go extinct. In the last decade three
species have gone extinct, with rates of extinction predicted to worsen
over the next two decades [threats to nature et al report]. Many
examples of opportunities to change our trajectory are coming to head
(@kearney2022, ). Australia also has the best research institutes in the
world dedicated to conservation research and thus punches significantly
above its weight in the research/identification/problem outputs.

JW: Summarise threatened species and ecosystems. The rates of species
endangerment and ecosystem decay is on the increase.

### Para 2 - Conservation crisis in Australia

Key papers: @whitelawBiodiversityDataPublic2021,
@williamsFutureRoleConservation2020a,

Much effort within conservation science has been dedicated to describing
the state of the biodiversity and species [DESCRIBE STATE]. And
identifying the proximate drivers of the change [UNDERSTAND MECHANISMS]
[political systems, ]. Lack of political motivation to balance human-led
impacts with equal or better commitments to preventing the driving
mechanisms and designating PAs [ESTABLISH SOURCE]. Additional to the
biodiversity crisis, there is a conservation crisis (coin phrase here)
in Australia. The broader community and leaders/representatives
designate a lack of attention, and cosnequently actions to the
biodiversity crisis [PROPOSE]. We coin this awareness gap as a
"social/political biodiversity crisis". This crisis manifests in that
insufficient attention to improving policy and providing the necessary
resources to conservation is occurring within Australia. Scientist, and
governance reviewers all identify these as significant constraints that
we need to break down to improve conservation outcomes
@australiannationalauditoffice2022, @samuel2020, @reside2019 .

Whilst, this has all been acknoledged, catalysing change within this
space has not been the focus of research or the community (beyond
eNGOs?). This may be because of traditional conservation science
research undervalues the production of social interventions. However,
Conservation science is a unique discipline in that it is relatively
connected with communities (citizen science), however there is a clear
opportunity for conservation sceicne to begin designing models that work
toward fixing the unerlying crisis of inaction (williams et al framework
of identifiying the drivers of extinction, maybe go back and dedicate a
whole para to this).

Use digital technology as a means of changing this to create (DESIGN).
Create a platform for this (IMPLEMENT).

JW: On top of a biodiversity crisis, there is a conservation crisis in
Australia. There conservation crisis is a lack of conservation attention
focussed at the biodiversity crisis. Key references National Audits,
Samuel reviews, review and perspective papers by scientists (Reside etc)
all show there is a lack of resources and action by leaders. Moreover,
and related, there are clear shortfalls in Australians knowing or
engaging with the issue. There is a clear need for new models in
conservation to catalyse change.

### Para 3 - Digital technology as an intervention for action

Key papers: Vrowmen APH paper,

"Individual and collective forms of political/social engagement are more
likely to occur via online means. @vromenPoliticalEngagementAustralian"

Australia has one of the highest usage rates of social media/smart phone
in the world @vromenPoliticalEngagementAustralian . Digital engagement
is an integral part of young people's lives and their political
interests. Although disagreements of the value of digital mediums as an
intervention for political engagement remain and the potential for
harm/good exist, many groups now use it as a means to communicate
stories/their activism. Given this rise in usage (and data
availability), opportunity for voices to be represented in politics is
greater than ever before [ASX blog post, vromen APH, ?]. Given the
adavancements in digital tech, there is a clear opportunity for concerns
around biodiversity to be communicated to the masses
@jolyMultimediaToolsApplications2018 .

In the current attention market (tech companies, other issues,
food/shelter responsibilities [ref ezra klein poddy]), we as a
conservation community cannot expect a disaffected public to pay
attention to an issue without a trade. Story-telling as a device for
social change has risen dramatically in the past decades as an effective
(?) means to broaden horizons. We must go beyond story-telling and offer
a service to capture attention and thus entertain through education, if
you must, to hook potential conservaitonists onto the issue. On the
other hand, the conservation movement (\$?) is dwarfed by corporate or
other interests (actually how much money does conservation attract
compared to other issues/disciplines?).

[PROPOSE/DESIGN]

Despite conservation being a mission-oriented discipline needing
advocates and social-media buy-in, it seems to bereft of social
media/digital platforms connecting people to the problem. Most
public-facing ecology/conservation initiatives are focused on the needs
of science (citizen-science) and the dissemination of results to
managers (shiny apps), with a lack of focus on inducting broader society
into the issue and encouraging adocacy for change.

In the last decade, social networks and digital technology has enabled
large communitties of citizen science to occur (e.g. inaturalist,
zooniverse, ebird, iSpot). Citizens have become increasingly aware of
the importance of monitoring biodiversity, the why behind it and have
leaped at the opportunity to assist @jolyMultimediaToolsApplications2018
. However, these efforts are important but not aimed at addressing the
'social conservation crisis'.

JW: Much hype around social media and apps to increase attention on
social issues. Climate example, health examples? Yet in conservation,
despite being a crisis discipline needing advocates, and wider society
buy-in, there seems bereft of these social media initiatives connecting
people to the problem. Most social media efforts in conservation and
ecology has been focussed on the needs of the science (key refs include
the papers already sent, we will need others) rather than disseminating
the finding to the audience to the general public to advocate for
change.

### Para 4 - Objective/aims

We deploy an ad-hoc issue-based digital-native organisation, called
'Threatened Australians' (TA) as means to experiment with adapting
conservation research/efforts as a social/digital intervention
[DESIGN/IMPLEMENT]. The Threatened Australians web-app operates at the
interface between data-driven/science serving (typically, citizen
science) platforms and political/social activism. It is the first
proactive attempt by a multidisciplinary team to provide an accessible
and engaging web application aimed at raising public attention around
the species crisis in Australia. The web-app was founded on aiming to
provide a resource that would encourage users to engage with the problem
and then offer easy means by which they can 'take action'. The primary
action of the web-app was facilitating communication with local federal
elected representatives via email, mimicking a tried and tested method
of political participation [vromen]. The web-app was timed for release
prior to the 2022 Australian federal election to raise attention as a
electoral issue, which was seen as lacking from the platforms of the
major parties (guardian article). We describe the processes involved in
the generation and deployment of the prototype. Second, we then provide
details of how we collected data around app usage and users perceptions.
Third, we synthesise this information to evaluate how it performed in
the Australian media/political landscape (EVALUATION/TEST) and across
electorates [TEST]. Finally, we examine models for improving robustness,
and future opportunities.

Such data is crucial to understanding the social interventions within
conservation.

JW: Here we provide a description of one of the first proactive attempts
by a multidisciplinary team to provide a web app aimed at raising public
attention around the species conservation crisis in Australia. Called
"Threatened Australians" (TA), its aim is to raise awareness of the
plight of threatened species, specifically there locations and threats.
It was linked to post code so to bring get local attention (Refs on need
for local action) and timed before the 2022 federal election so as to
bring attention specifically around federal politicians. Therefore the
key spatial boundaries used were federal electorates. We describe the
generation of the prototype and testing methods, public usage of the app
once realised and how this varied across different electorates.

## Methods

Key papers: @hendrieImpactMobilePhone2020,

Searching the app store for biodiverstiy apps...

### Para 1 - App design and development

Threatened Australians (TA) is available as a web app available at a
domain (threatened.org.au) through browser applications on mobile,
desktop, and tablet. The foundational idea of TA was to showcase the
threatened species ocuring in each 151 CEDs of Australia and encourage
users to take an action such as communicating with an elected
representative.

Once information was gathered (see Folder SX for methodology and data
information), planning and design of the app was undertaken. The design
of TA took on a bi-disciplinary approach with regular evaluations. The
development was led by a conservation science doctorate student with
counsel from a conservation science professor and an interaction design
expert. The team held an intial workshop with
scientists/campaigners/organisers from Australian conservation eNGOs
(WWF, TNC, Invasives Species Council, Birdlife, Bush Heritage, ACF,
Queensland Conservation Council) and multiple academics to receive
feedback on the idea and process. This initial workshop and subsequent
discussions confirmed the decision on the core function of the app. The
core goal of TA is to lead users to communicating with elected
representatives and connection with environmental groups/nongovernmental
organisations (eNGOs). During the design and development phase, the team
held meetings and communicated with a web developer to describe the
vision and reassess design decisions. Once developed to a functional
stage, the team held content and functionality testing workshops with
three eNGOs (WWF, BirdLife, ACF) and a design workshop with experts from
that field conducive to beta testing. Based on the feedback received,
the app was re-designed to accommodate requests and advice. Alpha
testing of the app did not occur.

Once users have provided their postcode, the application connects them
to the "set" of species within the associated CED. The first species
they are provided with is randomised, and the next, and so forth.

Sharing of the app functions.

We aggregated data from X source [how did we choose which ones?].

We entailed all data that intersected with Australia (offshore islands
etc) with no attention on specific bioregions/scales/urban-ness.

**Figure 1. an image of the front cover off the app showcasing its
features OR images of key pages and how users are to move through them
[@hendrieImpactMobilePhone2020]**

### Para 2 - App deployment and media

Accompanying the live version of the app was a media release, which was
sent out to Australian local and international media outlets.
Accompanying the release date (02/05/22) was a 'The Conversation"
article along with efforts being made via social media to promote TA.
The app creators personally messaged organisations and networks of
interest asking for air time on their platforms.

### Para 3 - Data collection

We generated data from two sources: passive analytics from usage and a
opt-in survey of perceived experience. Usage data was collected by
embedded Google Analytics 4 on the web app during a X week period from X
date to X date. Analytics data reflected usage based on the following X
indicators: (1) Page views and titles, (2) Acquisition of users, (3) X
something else, (4) demographics. Survey data came from a form
positioned in the header of the web app (Fig X) and was comprised of
four free-text answers and two likert [?] scale questions. We filtered
out the free-text questions (not enough volume, and kinda random
responses, Supplemetnary Material S\* [for full table of Q/A]), and
present the quantitative answers to the following questions: Q1 "Are you
going to take any actions as a result of using the web app?"; Q2 "The
information on this website was easy to understand:".

Information on media reach/other things from Dom Jarvis Media reach from
conversation article data

### Para 4 - Data processing and analysis

How app usage data was extracted.

Descriptive statistics of usage etc were used to...

All analyses were conducted using R and X packages.

Get shape file from world bank.

## Results

```{r setup, include=FALSE, fig.align='center', warning=FALSE, message=FALSE}
# knitr::opts_chunk$set(echo = TRUE, comment = "#")
# knitr::opts_knit$set(root.dir = "../")
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r libraries, include=FALSE}
library(tidyverse)
library(sf)
library(magrittr)
library(jsonlite)
library(units)
library(readxl)
library(stringr)
library(plotly)
library(forcats)
library(reactable)
library(readODS)
library(naniar)
library(viridis)
library(ggridges)
library(networkD3)
library(httpgd)
library(lubridate)
library(oz)
library(ozmaps)
library(mapdata)
library(maps)
library(tmap)
library(rmapshaper)
```

```{r import}
engag_title <- fromJSON(
  "data/clean_data/engag_title_clean.json"
)
species_title <- fromJSON(
  "data/clean_data/species_title_clean.json"
)
help_title <- fromJSON(
  "data/clean_data/help_title_clean.json"
)
acquis_time <- fromJSON(
  "data/clean_data/acquis_time_clean.json"
) %>%
  mutate(
    date = as.Date(date)
  )
acquis_traffic <- fromJSON(
  "data/clean_data/acquis_traffic_clean.json"
)
demo_country <- fromJSON(
  "data/clean_data/demo_country_clean.json"
)
demo_region <- fromJSON(
  "data/clean_data/demo_region_clean.json"
)
demo_city_geom_clean <- st_read(
  "data/clean_data/demo_city_geom_clean.gpkg"
)
links_domain <- fromJSON(
  "data/clean_data/links_domain_clean.json"
)
links_email <- fromJSON(
  "data/clean_data/links_email_clean.json"
)
feedback <- fromJSON(
  "data/clean_data/feedback_clean.json"
)
elects_demo <- fromJSON(
  "data/clean_data/demo_clean.json"
)
media <- fromJSON(
  "data/clean_data/media_clean.json"
) %>%
  mutate(
    published_date_time_AEST = ymd_hms(published_date_time_AEST)
  )
elects <- st_read(
  "data/clean_data/elect_clean.gpkg"
) %>% 
  ms_simplify(.,
  keep = 0.01,
  keep_shape = TRUE
) %>%
  st_make_valid()
```

### Para 1 - Description of users and timing

Released XXX. #% of users prior to election day. Over the course of X
days up to federal election, \# Number of people and hits over time
(Figure 2).

**Figure 2. a time series of usage, including key media events** - How
did our traffic acquisition change over time, drawn next to key media
events?

```{r, fig_width=20, fig_height=20, fig_caption="Figure 2. a time series of usage, including key media events"}
# TODO: add in key media events, maybe by joining the two datasets, or manual add-in as per Chang 2018
acquis_time %>%
  select(
    !where(
      ~ is.numeric(.) &&
        sum(.) <= 200
    )
  ) %>%
  pivot_longer(
    cols = -date,
  ) %>%
  rename(
    session_source = name,
    users = value
  ) %>%
  filter(
    between(
      date,
      as.Date("2022-05-02"),
      as.Date("2022-05-30")
    )
  ) %>%
  ggplot(.) +
  aes(
    x = date,
    y = users,
    fill = session_source
  ) +
  geom_area() +
  # geom_text(
  #   data = media,
  #   aes(
  #     x = published_date_time_AEST, #TODO: maybe this has something to do with the ymd format I got it in
  #     label = platform_name,
  #     y = 0
  #     )
  #   ) +
  scale_fill_viridis(
    option = "H",
    discrete = TRUE
  ) +
  ylim(0, 5400) +
  theme_minimal() +
  labs(
    x = NULL,
    y = "Number of users",
    fill = "Acquisition"
  ) +
  annotate("segment", x = as.Date("2022-05-03"), xend = as.Date("2022-05-04"), y = 2500, yend = 2700) + 
  annotate("text", x = as.Date("2022-05-05"), y = 2850, label = "The Conversation", size = 2.5) +
  # annotate("segment", x = as.Date("2022-05-15"), xend = as.Date("2022-05-15"), y = 5150, yend = 5300) +
  annotate("text", x = as.Date("2022-05-15"), y = 5300, label = "ABC Science", size = 2.5) +
  annotate("segment", x = as.Date("2022-05-24"), xend = as.Date("2022-05-24"), y = 500, yend = 750) + 
  annotate("text", x = as.Date("2022-05-24"), y = 900, label = "Behind the News", size = 2.5)
  
```

### Para 2 - How patterns differentiate

Number of electorates and density of hits across space. Describe
patterns

**Figure 3. A map of Australia with hits.** - How do the locations our
users come from differentiate?

A)  Could do one chart of sum species page view per electorate then B)
    sum how to help page per electorate

```{r, fig_caption="Figure 3. A map of Australia with hits on a per city basis"}
species_title_geom <- elects %>% 
  full_join(., species_title) %>% 
  group_by(electorate, state_territory, state_territory_abbrev, demographic_class) %>% 
  summarise(
    active_users = sum(active_users),
    engaged_sessions = sum(engaged_sessions),
    event_count = sum(event_count)
  )

# TODO: started the dorl 
species_title_geom_dorl_weight <- species_title_geom %>%
  st_transform(
    3112
  ) %>%
  cartogram_dorling(
    weight = "event_count",
    k = 0.3
    # m_weight = 1
  )

tm_shape(species_title_geom) +
  tm_polygons() +
  tm_bubbles(size = "event_count")
# going per polygon? LOL

help_title_geom <- elects %>% 
  full_join(., help_title)
tm_shape(elects) +
  tm_polygons() +
  tm_shape(help_title_geom) +
  tm_bubbles(size = "event_count")
# but not here??


spec.per.elect.unique.dorl.weight <- spec.per.elect.counts.summary %>%
  st_transform(
    3112
  ) %>%
  cartogram_dorling(
    weight = "total_unique_species",
    k = 0.3
    # m_weight = 1
  )

#### tmap ####

spec.per.elect.unique.spec.dorl <-
tm_shape(
  elect,
  bbox = st_bbox(c(
    xmin = 112.929704,
    ymin = -43.627943,
    xmax = 155.629864,
    ymax = -9.115517
    )
  )
  ) +
  tm_borders(
    "white",
    lwd = 1
  ) +
  tm_fill(
    "#d8d8d8"
  ) +
  tm_shape(
    spec.per.elect.unique.dorl.weight,
    bbox = st_bbox(c(
      xmin = 112.929704,
      ymin = -43.627943,
      xmax = 155.629864,
      ymax = -9.115517
    )
  )
  ) +
  tm_fill(
    "total_unique_species",
    style = "jenks",
    # style = "cont",
    title = "Threatened species",
    palette = "-inferno"
  ) +
  tm_text(
    "electorate_abbrev",
    size = "AREA"
  ) +
  tm_borders(
    alpha = 0.6
  ) +
  # tm_legend(
    # legend.position = c("left", "top")
    # title.position = c("left", "bottom")
  # ) +
  # tm_compass(position = c("left", "bottom")) +
  # tm_scale_bar(
  #   position = c("left", "bottom"),
  #   width = 0.2
  # ) +
  tm_layout(
    frame = FALSE,
    # legend.width = 1,
    # legend.height = 0.8,
    # legend.title.size = 1.5
    # legend.outside = TRUE
  )


aus <- ozmap(x = "country")
demo_city_geom_clean <- demo_city_geom_clean %>% 
  filter(!city == "Unknown")
# TODO: stuck on contours, maybe need to rasterise?
# tmap option
tm_shape(aus) +
  tm_polygons() +
  tm_shape(demo_city_geom_clean) +
  tm_bubbles(size = "active_users")
# ggplot option
ggplot(aus) +
  geom_sf() +
  geom_sf(data = demo_city_geom_clean) +
  geom_density2d(
    data = demo_city_geom_clean, 
    aes(fill = active_users),
    alpha = .5,
    geom = "polygon"
  )



```

Number of species investigated and most popular species were XXX.

**Figure 4. a map of the locations of the top 5-10 species and how many
hits they had, and a photo of each of them** - How do the top 5-10
[meaningful measure] differentiate?

We have `r names(species_title)[7:12]` to choose from, i.e. 6
measurements. It seems every one of them is influenced by the
sitribution of the species and how often users are likely to see it. We
could maybe get around this by doing some sort of normalisation, but
how?

```{r, fig_caption="blah"}
exp <- species_title %>%
  group_by(vernacular_name) %>%
  summarise(sum_active_users = sum(active_users))
```

Optional: **Figure 5. Sankey of movement throughout platform** - How did
our users travel through the app? [user journey] - How did they differ
at the electorate/species level?

## Discussion

### Para 1 - Summary

JW: We provide the first app that attempts to put species and threat
information on Australia's threatened species into the hands of the
public. In the space of a month, -- users, across all federal
electorates engage with it showing there was widespread interest. There
was also media interest, with a ABC Science digital story and
`r length(media$platform_type[media$platform_type=="radio"])` radio
interviews conducted on local stations around the country. The appetite
seemed there.

### Para 2 - How to do it more robustly

"here is what kind of scientific impact can be achieved on a small budget, here is how you can learn from our approach"

JW: This is a first attempt and done via just two rounds of review by
NGOS. A more systematic way of doing this would be Y, X, and Z (use the
theory in the paper that was shared to describe how to more robustly do
it but also need to refer to other non biodiversity literature about
what is working and what we could do better.

### Para 3 - Further customisation/targeting is needed

Modern organisations (corporate and NFP) have heavily invested or begun
with database management/linkage that enable optimisation of services to
the preferences of an individual. We did none of this.

Particularly in conservation, organisations guard current and
prospective supporters jealously. This insecurity and dedication to
their own mission disables their abilities to coordinate together. As
cooperation is an establish foundation of human achievement, we cannot
expect to bend the curve of species loss/induct further followers into
the mission-oriented focus of conservation if this continues.
Furthermore, this causes

JW: Further work is needed to catalyse not just linking the threat to
place, but the actual actions needed to be activated in the place. This
type of connection between people and actions is likely critical pathway
to action.

Constantly evolving socio-political context

### Para 4 - Education opportunities

JW: There is also significant opportunities with the education dynamics
and children. Children in schools increasingly use social media as part
of their learning packages (refs) and a app that links education to
awareness and advocacy is clearly missing.

### Para 5 - Conclusion

JW: Our efforts are a first attempt to get over a gap in the
conservation space around making people aware of the challenges
biodiversity faces and what they can do about it. It takes advantage of
significant gains in science over the past decade documenting ranges and
threats to species (cite NESP 7.7) and tries to generate information in
a accessible to the general public. More is needed to get action etc and
we have attempted to provide one model to bridge the knowledge-advocacy
gap.

The progression of conservation science from identifying the causes to
designing and implementing the solutions has been explained in this
study. Whilst the key step of evaluating the effectiveness of the
intervention has not been undertaken
(@williamsFutureRoleConservation2020a). This represents a key step for
improvement in the kind of intervention we are examining here today.

### Para X - Caveats

-   Didn't organise anything prior to app release (lol)
-   Science underlying digital tech conservation apps varies alot, our
    base data is rubbish

## Supporting information

## Acknowledgements

The authors would like to recognise...

The authors declare no conflicts of interest.

This work was funded by the University of Queensland Centre for
Biodiversity and Conservation Science, Green Fire Science laboratory,
and the Queensland University of Technology Media centre thing.

## Data

All data used to produce the Threatened Australians database is
available on GitHub.

Raw, clean, and analysed data produced by Threatened Australians is
available on UQ eSpace
