---
title: "Leveraging digital technology to catalyze public will for saving biodiversity"
format:
  docx:
    reference-doc: /home/gareth/science/admin/bibliography/word_templates/research_article_template.docx
    fig-width: 6
    fig-height: 6
bibliography: /home/gareth/science/projects/electoral/threatened_australians_research/lenovo_pop_os_my_library_spec.bib
crossref:
  ref-hyperlink: true
execute:
  echo: false
  warning: false

---

```{r}
#| include: false

pacman::p_load(tidyverse, httpgd, here, ggrepel)
hgd()
here::here()
here::i_am("writing/TA_theory/Kindler_2023_digital_tech_biodiv.qmd")
```

# Article admin

## Alternative titles

- How to overcome lack of public will for saving biodiversity using digital technology

## Authors

Gareth S. Kindler^1,2^\*, Nick Kelly^3^, Tim Carden^3^, James E.M. Watson^1,2^

## Affiliations

1. Centre for Biodiversity and Conservation Science, The University of Queensland, St Lucia 4072, Australia
2. School of Earth and Environmental Sciences, The University of Queensland, St Lucia 4072, Australia
3. School of Design, Queensland University of Technology, Brisbane 4059, Australia
\ \

\***Corresponding author email**: [g.kindler\@uq.edu.au](mailto:g.kindler@uq.edu.au){.email} \ \

**Keywords**: biodiversity, digital technology, conservation, design, political will, public will, persuasion, social influence, pro-environmental behaviour

## Draft target journal - Nature Ecology & Evolution as a Perspective

Nature EE's main points in response to our article pitch:

- Interested in discussing biodiversity enhancement as an election issue
- Interested in the linking of elected officials to biodiversity crisis
- Concern over narrow focus on Threatened Australians experience and representative democracies (many systems exist in the world)
- Suggest to collaborate with co-authors in other locations
- Showcase multiple case studies (2 or 3) for broader perspective
- General framing: emphasize making biodiversity an election issue, that the public are aware they can contribute to decisions about
- Discuss why it isn't a prevalent issue yet
- Mention successful examples (e.g., Costa Rica)

## Target audience

- From [NEE's aims page](https://www.nature.com/natecolevol/aims), they are interested in:
  - How life of earth is maintained, how we should conserve it
  - The implications of ecology stretch beyond biology into environmental science and the grand challenges facing society
  - Relevant parts of the social sciences
  - And is a place where all researchers and policymakers interested in all aspects of life's diversity can... learn about advances and discuss topical issues

<!--

[Submission guidelines](https://www.nature.com/natecolevol/content)

Perspective

A Perspective is a format for scholarly reviews and discussions of the primary research literature that are too technical for a Comment (see below) but do not meet the criteria for a Review—either because the scope is too narrow, or because the author is advocating a controversial position or a speculative hypothesis or discussing work primarily from one group. Two reviews advocating opposite sides in a research controversy are normally published as Perspectives.

Format

    Main text – 3,000-4,000 words,
    4-6 display items (figures, tables or boxes).
    No footnotes.
    References – up to 100 (exceptions are possible in special cases).
    Reviews include received/accepted dates.
    Reviews are peer reviewed.

Comment examples: [1](https://www.nature.com/articles/s41559-020-01285-z), [2](https://www.nature.com/articles/s41559-022-01976-9), [3](https://www.nature.com/articles/s41559-020-01285-z)

[Perspective examples](https://www.nature.com/natecolevol/reviews-and-analysis?searchType=journalSearch&sort=PubDate&type=perspective&page=1): [1](https://www.nature.com/articles/s41559-020-1230-6), [2](https://www.nature.com/articles/s41559-020-1275-6), [3](https://www.nature.com/articles/s41559-020-1253-z)

-->

## Potential co-authors/reviewers

1. [Ariadne Vromen](https://crawford.anu.edu.au/people/academic/ariadne-vromen) (ANU)
2. [Mitchell Whitelaw](https://researchers.anu.edu.au/researchers/whitelaw-m) (ANU)
3. [Jean-Baptiste Gouyon](https://www.ucl.ac.uk/sts/people/dr-jean-baptiste-gouyon) (UCL)
4. [Jagadish Thaker](https://communication-arts.uq.edu.au/profile/7112/jagadish-thaker) (UQ)
5. [Carl DiSalvo](https://www.cc.gatech.edu/people/carl-disalvo) (Georgia Tech)
6. People from that UQ Centre for Policy Futures?
7. [Katherine Cullerton](https://public-health.uq.edu.au/profile/3786/katherine-cullerton) (UQ Public Health) - I had coffee with her once

<!--
## Search terms

Public will:

- "policy backing" OR "political will" OR "public will" OR "decision-maker consensus" OR "support from key stakeholders" OR "political endorsement" OR "policy buy-in" OR "leadership agreement" OR "commitment from authorities" OR "unified support" OR "public understanding" OR "public awareness" OR "behaviour change" OR "effective policy" \ \
-->

## Current caveats/questions/problems to address

**General**

- Some sections of the manuscript are further developed (Intro) than others (Sections 1, 2, 3). I've intentionally done this instead of reducing everything to draft structure as I'm operating on the assumption that more info is better for now
- Our focus on what we are trying to catalyze is unclear
  - Right now I chop and change between raising awareness, public will, and political will. They are all related and overlapping but it might be helpful to choose one. For now I'm going to lump them all together under 'PPW'
- Defining the problem
  - This is related to above. We haven't sufficiently defined the problem that we are addressing using data/studies. For example, is it that the public don't know the extent of the crisis (a commonly spouted problem), or is it that they know but are not motivated to exercise their democratic accountability powers over it, or rather something else. We can make some assumptions based on the current outcomes (poor progress). We could probably get by with saying that broad awareness is always desirable (I have seen this done in the context of CC).
  - There are advantages of being more specific though. If the report ACF shared with us was public this would solve this problem. Maybe [Chris Sandbrooke's Future of Conservation](https://futureconservation.org/) (global) or the new Biodiversity Concerns publication (AUS only) has some data that we could repackage to better define the problem such as by saying that "X% are unaware of the biodiversity crisis and the consequences" and "X% want more government action but don't know how/what to do to encourage it"

**Section specific**

- Introduction
  - Currently a vaguely connected series of relevant paragraphs. This will be edited/changed to suit our Sections when we have agreed on a structure. I think we can all agree that devising the Intro, given how short they are in Perspectives is low hanging fruit but can't be picked til the rest is
- Section 1
  - Figure 1 (Quadrant)
    - Words used to describe the quadrants are very much drafts. I am aware they don't make sense/capture what we are trying to say
    - The plotting of projects along the axes means that we are making precise subjective judgements about where we perceive them to be (how political or what their motives are). When we consider intention or outcome this is complicated further as there will almost definitely be discrepancies. Maybe best to focus on what the intervention's 'intention' is then.
    - Our options appear to be that we can:
      1. Remove them from the graph and discuss them as examples when referring to each quadrant
      2. Keep them in the graph
      3. Something else I haven't thought of?
  - I'm conceiving this section as a walk through of Figure 1 answering those questions I've put in bold with 'Main question' beforehand
- Section 2
  - Do you guys think this Section is relevant for our article?
  - Section 2 could well be expanded or shortened. Our options seem to be:
    1. Make it comprise the second half of the article (as is)
    2. Consolidate it into Section 1
    3. Drop and focus on Section 1 being all of the body
- Section 3
  - Seems like a necessary section, lmk if not
- Appendix
  - Figure 2 (Conceptual diagram of policy change)
    - Is something like this useful for our purposes?

\newpage

# Abstract

Despite its significance, biodiversity loss is the worse off and less known global environmental threat. Because it is a wicked and collective action problem, it requires system-level change to make progress. Yet, the movement has not been able to convince institutions to act at a sufficient scale to curb declines and extinctions. The interventions that conservation has developed have fallen short on engaging the broader public, and building political and public will. In this perspective, we explore how the movement has neglected designing broadly palatable interventions that explicitly create these building blocks for policy change, the draw cards of biodiversity in this context, and case studies of interventions. We finish by proposing a research agenda aimed at convincing governments to act.

# Introduction

Biodiversity is the worse off and less known global environmental threat [@steffenwillPlanetaryBoundariesGuiding2015]. As well as being affected directly, it is representative of all other Earth system stressors [@lovejoyHabitableEarthProtecting2019]. Current projections for biodiversity are expecting the situation to get much worse.

Conservation science has focused on describing the state of the problem such as identifying species and ecosystems that are at risk [@williamsFutureRoleConservation2020]. This has left behind a poor track record of translating effort into the desired change, or rather a field plagued by incrementalism [@evansReconceptualizingRoleScience2021; @toomeyNavigatingSpaceResearch2017]. Ambitions to save all species have been replaced by triage of the most threatened species and ecosystems.

Yet, the crisis is driven by the activities of humans and the resources that they consume. This demands a social response. Whilst changes in the activities of individuals can help, system-level changes are required to be transformative [@leiserowitz17BuildingPublic2019]. Attempts at system-level change through conventional policy solutions aimed at limiting the activities of humans have largely failed [@schmitzSustainingHumansNature2019; @ipbesSummaryPolicymakersGlobal2019]. The lack of mainstreaming of biodiversity into public policy has ensured the failure of these attempts [@rounsevellBiodiversityTargetBased2020; @intergovernmentalscience-policyplatformonbiodiversityandecosystemservicesRegionalAssessmentReport2018].

Mainstreaming/system-level/transformative change is only surmountable with government interventions such as policy solutions. Conserving nature requires collective action due to its inherently complex and wicked nature. Unlike well-defined, solvable problems, conserving nature involves a myriad of interdependent factors, including ecological, social, economic, and political dimensions, making it complex and difficult to tackle.

Commitment to delivering transformative policy will require broad support. Policy is more likely to be enacted if there is a wide range of support from politicians [@eylerPatternsPredictorsEnactment2012]. It is therefore important to engage the public and decision-makers in the crisis [@roseMajorBarriersEvidenceinformed2018; @raileDefiningPublicWill2014; @postDefiningPoliticalWill2010]. Raising awareness is also a precursor to "problem recognition and solution seeking" [@goldbergShiftingRepublicanViews2021; @goldbergLeveragingSocialScience2020], which are key foundations in cultivating public and political will (PPW) for policy change.

PPW are distinct concepts that share dimensions. Public will is the public's desire for change, whereas political will is the political establishment's desire for change. Both are required to achieve transformative change. Public will is a necessary precursor to political will, but not sufficient. Political will is required to enact policy change. Political will is also required to sustain public will.

The conservation movement needs to mirror the respect/prioritisation found among scientists in the widespread awareness and understanding of global populations. An informed and caring populace can actively contribute to biodiversity preservation and fortify our collective resilience against impending environmental challenges.

Information and Communication Technology offers an increasingly promising avenue to achieve these goals. It reaches a broad audience and is responsive to design changes, making it a cost-effective solution. As people, especially younger generations, increasingly embrace digital worlds and expression, ICT has become an integral part of our lives and political interests [@vromenPoliticalEngagementAustralian]. Political engagement is more likely to occur through online means, creating an unprecedented opportunity for voices to be represented, thanks to the widespread usage and accessibility of ICT. [@vromenPoliticalEngagementAustralian]. There is the "solutionism" trap that comes with technology and the development of applications. Our argument is that we should use a technological solution to affect what is seen as politically untenable or impractical.

In this perspective, we outline how the movement has neglected designing broadly palatable interventions that explicitly lead to policy change. Followed by the draw cards of biodiversity in this context, and case studies of interventions. We finish by proposing a research agenda aimed at convincing governments to act.

# Section 1: Social interventions with a political valence

The conservation movement has produced a diverse spectrum of ICT interventions, each with unique contributions to the mission. Among these are four distinct categories (@fig-quadrant). First, we have science-serving projects like citizen science initiatives and the Atlas of Living Australia, which primarily disseminate or accumulate scientific data. Second, we see ecologically-minded projects, such as NGO and government agency campaigns, which despite their public outreach, may fall short due to their eco-centric and audience focus. Third, we find layperson-charismatic initiatives, including National Geographic and documentaries, that offer high accessibility for the general public but do not promote actionable political change. Lastly, we identify politically-charged projects, like Extinction Rebellion and Threatened Australians, which aim to transform public interest into substantial political action. It is the final category, the politically-charged projects, that appears to be under-explored and under-utilized in the context of conservation efforts. Harnessing the potential of this category could mark a crucial turning point in driving socially impactful conservation action through explicitly building PPW.

## Science-serving

- **Why have science-serving initiatives not delivered on PPW?**

Social networks and digital technology has enabled large communities of citizen science to occur (e.g. iNaturalist, Zooniverse, eBird, iSpot). In parallel, science has created tools with similar attributes but with a different target audience such as conservation practitioners or scientists themselves. These tools have helped the public become increasingly aware of the importance of monitoring biodiversity, the why behind it and have leaped at the opportunity to assist [@jolyMultimediaToolsApplications2018]. We propose these efforts are targeting or only attracting the ecologically minded [REF], and thus a insignificant amount of the general population [@whitelawBiodiversityDataPublic2021]. These public-facing ecology initiatives are focused on the needs of science, with little to no focus on inducting broader society into the issue.

## Ecologically-targeted

- **Why have ecologically-targeted (NGO and government) initiatives not delivered on PPW?**
- Competition and collaboration - orgs are disorganised, have diverse agendas, guard supporters jealously, impeding coordination. Why is this unique to conservation? Is it because we have objects (a woylie) in physical space with monetary demands?
- Collective action problems require a critical mass of people to act, this mass can be smaller if the other people are not affected by the change/against it. @goldbergShiftingRepublicanViews2021 targeted republicans as a group/civic that is important to achieving robust climate policy. They had no intention to convince everyone, they focused on a subset of the population who were prone to blocking further action.

## Layperson-charismatic

- **Why have layperson-charismatic initiatives not delivered on PPW?**
- Celebrity environmentalists such as Attenborough, Irwin, Goodall have all successfully used biodiversity content in their platforms and have been making 'calls to action' for change. Why has this been ineffective?

## Politically-charged

- **Why will politically-charged initiatives deliver on PPW? Or why have they?**
- We think of democracy (the political process) as having rationality/order marred by structured decision-making processes. This is not the case, it is contentious, adversarial, full of disagreement, conflict and confrontation.
- An agonistic/agonistic pluralism grounding accepts that democracy is intrinsically contentious
- Carl DiSalvo says the "most basic purpose of adversarial design is to make these spaces of confrontation and provide resources and opportunities for others to participate in contestation"
- Beyond that, DiSalvo claims that there are two types of adversarial design. There is design to support and improve the mechanisms and procedures of governance, and then there is to create contestation and confrontation
- Social interventions that aim to catalyse public will for policy change are inherently adversarial and comprise both types of adversarial design, as seen with Threatened Australians. TA was was about raising awareness so others can be aware of 'the problem/fight/mission' and join if they are motivated whilst simultaneously supporting civics and engagement in political process

```{r}
#| label: fig-quadrant
#| fig-cap: "The four quadrants of Information and Communication Technology (ICT) social interventions"

source(here("R/fig_quadrant.R"))
ggplot_quadrant(quad_data)
ggplot_save("quadrant", ggplot_quadrant(quad_data))
```

# Section 2: Initiating and sustaining engagement

## Biodiversity in a competitive information environment

- Why is biodiversity unique in a competitive information environment?
- How do these attributes put biodiversity at an advantage?
- How do these advantages compare to other mission-driven disciplines (climate/agricultural/public health)?
- How can these advantages be leveraged to build PPW?

To capture the attention of individuals, conservation needs to compete for attention in an existing information environment/attention market. Biodiversity has unique practical, emotional, and philosophical advantages that can assist in attracting attention to the mission of saving it. These are ecosystem services, cultural and spiritual values, aesthetic value, discovery (scientific and non-scientific), creative inspiration, mesmerising adaptations/behaviours (biolumin, traits), moral and ethical considerations, psychological well-being. There is risk of these advantages being fetishized/romanticised to the public, but it affords the opportunity to showcase what is at stake.

Organisations often operate on the premise that they has something to "sell" or that the consumer "wants". In the context of raising awareness to save biodiversity, people are needed to achieve this mission because the human and non-human are intertwined in the wicked problem context. Not as a consumer, they are a partner (or a kind of consumer of nature?). This is a fundamental difference that needs to be understood and capitalised on in the design of PPW interventions.

## Mental models/schemas

- Why are mental models and schemas relevant to generating lasting engagement? Such as threats to self and nature, interconnected systems, way of life, regenerative agriculture, indigenous histories [@goldbergLeveragingSocialScience2020]

## Tailoring experiences

- How can we use research on human preferences to reach broader audiences and be more impactful?
- There are many messages to spout such as deepening the public's understanding of nature, the crisis, the importance of functional landscapes
- Importance of drawing upon theory, field, and lab experiment-based guidelines for messaging/building awareness
- Biodiversity awareness building needs to be careful to not fall into either tokenism or exploitation

Publics are motivated by different reasons to take action. "Building public awareness of biodiversity requires an understanding of the publics within a population"[@leiserowitz17BuildingPublic2019] [Remember the ACF report they shared with us, this is what that did]. Biodiversity must compete against well-resourced legacy and new media companies (TikTok/Meta). These modern organisations (FP and NFP) have heavily invested in tailoring the experience of their users to individual preferences (e.g. recommender algorithms), to increase engagement. Research into human preferences has revealed information on biodiversity that would make tailoring digital experiences to boost engagement feasible [@thommesWhatInstagramCan2021; @tribotIntegratingAestheticValue2018]. Tailored engagement strategies are needed to reach different publics and sustain their involvement.

## Persuasion

- How do we persuade people to go from unaware/uncaring to aware/caring to motivated/active?
- Recently, @goldbergLeveragingSocialScience2020 reported the social science research integral to generate lasting engagement change with climate change solutions

# Section 3: Bringing it together

- What are the key points we have made?
- What are the caveats or considerations?

# Appendix

![Conceptual diagram of the interactions between public and political will and policy change](`r here("fig/ppw_enablers_barriers.pdf")`){#fig-ppw-enablers-barriers}

# Supporting information

All data is on a public, permanent repository found here.
All materials are available on a public repository found here.

# Acknowledgements

G.S.K, N.K, and J.E.M.W conceived of and designed the research. G.S.K drafted the work. G.S.K, N.K, T.C and J.E.M.W worked on the development of discussed topics. All authors contributed to the article with substantial revisions and approved the submitted version.

The authors declare no conflicts of interest.

This work was funded by the University of Queensland Centre for Biodiversity and Conservation Science, Green Fire Science laboratory, and the Queensland University of Technology School of Design.

# References