---
title: Cleaning of the analytics and feedback data<br><h2>Threatened Australians (threatened.org.au)
author: Gareth Kindler<br>The University of Queensland, Australia
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    html_document:
        code_folding: show
---

```{r setup, include=FALSE, fig.align='center', warning=FALSE, message=FALSE}
# knitr::opts_chunk$set(echo = TRUE, comment = "#")
# knitr::opts_knit$set(root.dir = "../")
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r Libraries, message=FALSE}
library(tidyverse)
library(sf)
library(magrittr)
library(jsonlite)
library(units)
library(readxl)
library(stringr)
library(plotly)
# library(RGoogleAnalytics)
library(forcats)
library(reactable)
library(readODS)
library(naniar)
library(viridis)
library(ggridges)
library(networkD3)
library(maps)
library(vroom)
library(data.table)
```

# Import data

## Feedback

The other type of data is feedback data - this is a single data frame with the answers to the feedback survey

Export it by copy and pasting the data into a new spreadsheet.

```{r read-feedback}
feedback <- read_ods(
  "data/raw_data/TA_feedback_output/21-to-22-06-16_feedback.ods"
)
```

## Media appearances

```{r}
media <- read_ods(
  "data/raw_data/media.ods"
)
```

# Clean

## Feedback

Cleaning the feedback data was pretty easy besides not being able to figure out why I couldn't convert the blankspaces to `NA`s. Felt like I tried everything, so I gave up. If I want to pick this back up, start with trying to extract the regex characters of the missing values as doing it the other way around wasn't working. A work around for the multiple choice questions was to set the levels for the factor and anything left over (the blanks) get converted to `NA`s.

```{r feedback-clean}
feedback_clean <- feedback %>%
  rename(
    submission_id = submissionid,
    key_takeaway = answer1,
    actions_result = answer2,
    actions_be = answer3,
    easy_to_understand = answer4,
    improvments_suggestions = answer5,
    anything_else = answer6
  ) %>%
  filter(
    !str_detect(
      key_takeaway,
      "test|Test|выплат|готовы")
  ) %>%
  filter(
    !submission_id %in% c(1:11, 47:49)
  ) %T>%
  write_json(
    "data/clean_data/feedback_clean.json"
  )
```

After removing test submissions of the feedback form and some dodgy answers, we went from `r dim(feedback)[1]` to `r dim(feedback_clean)[1]`.

## Media

```{r}
media_clean <- media %T>%
  write_json(
    "data/clean_data/media_clean.json"
  )
```
```